MinCall --- MinION od raw signala do genoma

### 2017-05-11 paper/main.tex
Update
Updated abstract

\begin{sazetak}
MinION je citac DNA najnovije generacije. Oxford Nanopore Technologies ga je u svrhu prvog portabilnog DNA sekvencora. Proizvodi vrlo duga ocitanja (duljine i preko 100 kBp), doduse pod cijelu preciznosti. U ovom radu prezentiramo altrenativnu metodu ocitanja DNA koristenjem convolucijskim neuronskih mreza (CNN) umjesto dosadasnjeg pristupa s HMM ili RNN. Takoder cijeli postupak je istreniran od izvornog signala do ocitanih nukleoida, bez ovisnosti o umjetnoj segregaciji na dogadaje.

@@

\begin{abstract}
    The MinION device by Oxford Nanopore is the first portable sequencing device. MinION is able to produce very long reads (reads over 100 kBp were reported), however, it suffers from high sequencing error rate. In this paper, we show novel model basecaller which basecalles 1D reads using gated residual convolutional neural networks (CNN) combined with Connectionist Temporal Classification (CTC) learning approach. Its performance is on par with state of the art and offers possible future improvement.

    The whole process is trained end-to-end from the raw signal on R9 sequencing pore and once trained it's independent of Metrichor.
\keywords{Basecaller, MinION, R9, CNN, CTC, Next generation sequecing}
\end{abstract}

@@

@@

In out concrete $x$ aplication $\Sigma = \{\text{A G T C A' G' T' A'}\}$

@@

>>>>

\begin{sazetak}
    U ovom znanstvenom clanku istenirali smo model dubokog ucenja za Oxford Nanopore Tehnology's MinION portabilni DNK sekvencor. MinION potencijal je ogroman, i zbog njegove male velicina i portabilnosti, kao i dugih ocitanja senzora (ocitanja > 100kbP su primjecena). Nazalost pati od vecih gresaka u usporedbi s drugim modelima.

    Kreirali smo nezavisni model pocevsi od izvornih podataka i koristili konvolucijske neuronske mreze (CNN) s CTC gubitkom te postigli mjerljive perfomanse.

@@

\begin{abstract}
    In this study we create end2end deep learning model for Oxford Nanopore Technology's MinION portable DNA sequencing device. The device's potential is enormous, both due to it's small size and portability and long read (over 100kBp reported). However is suffers with higher error rate compared to other methods.

    We've created independant model starting with raw data and used Convolutional Neural Network (CNN) combined with Connectionist Temporal Classification (CTC) loss for measurable performance.
\keywords{Basecaller, MinION, R9, CNN, CTC, Next generation sequecing}
\end{abstract}

@@

The exact error rate metric is unreliable since multiple pipline tools could be the issue. First the sample is prepared, hopefully uncontaminated and matching reference genome as close as possile, then sequenced using the MinION device obtaining raw data. Next our model (or other groups ones) come along, basecall the sequence. To evalute error rate metric basecalled read is aligned to reference using aligners with their own erros/biases, mostly commonly used BWA-MEM and Graphmap.

@@

In out concrete $x$ application $\Sigma = \{\text{A G T C A' G' T' A'}\}$

@@

\chapter{Acknowledgments}

TODO.
